### 2017-11-21 manuscript.rmd
rearranged supplementary info, removed some extras

@@

### Supplementary Analyses

```{r char_evo_setup, include=F, echo=F}

#Load data
read.csv('character_coding/main_characters.csv', header = T, sep = ',') -> cdata
cdata$Species = str_replace_all(cdata$Species, "sp ", "sp")
rownames(cdata) = cdata$Species

#####The Tree

```{r ultratree, echo=F, message=F, warning=F}
par(ask=F)
par(mfrow=c(1,1))
plotTree(ultratree, lwd = 4)

```
```{r simmap_making, echo=F, message=F, warning=F}
#Stochastic trait mapping:
#Make simmap objects
#print("SIMMAP Sex Distribution")
make.simmap(ultratree, sex, nsim = 100, message=FALSE) -> sex_sim
#print("SIMMAP Palpons")
make.simmap(ultratree, palpons, nsim = 100, message=FALSE) -> palpons_sim
#print("SIMMAP Tentilla")
make.simmap(ultratree, tentilla, nsim = 100, message=FALSE) -> tentilla_sim
#print("SIMMAP Tentilla2")
make.simmap(ultratree, tentilla2, nsim = 100, message=FALSE) -> tentilla_sim2
#print("SIMMAP Nectosome")
make.simmap(ultratree, nectosome, nsim = 100, message=FALSE) -> nectosome_sim
#print("SIMMAP Nectosome Position")
make.simmap(ultratree, necto_pos, nsim = 100, message=FALSE) -> necto_pos_sim
#print("SIMMAP Nectophore Types")
make.simmap(ultratree, necto_types, nsim = 100, message=FALSE) -> necto_types_sim
#print("SIMMAP Bracts")
make.simmap(ultratree, bracts, nsim = 100) -> bracts_sim
#print("SIMMAP Habitat")
make.simmap(ultratree, vrange, nsim = 100, message=FALSE) -> vrange_sim

@@

#####SIMMAP Position of Nectosome
```{r nectosome_position, echo=F, message=F, warning=F}
#SIMMAP NECTOSOME POSITION
plotTree(ultratree, lwd = 4)
colors = c("black", "red", "green")
names(colors) = c("Dorsal", "No nectosome", "Ventral")
necto_pos_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(necto_pos_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
```

@@

#####SIMMAP Nectophore number and types
```{r nectophore_number_types, echo=F, message=F, warning=F}
#Simmap NECTOPHORE TYPES
plotTree(ultratree, lwd = 4)
colors = c("black", "red", "green", "blue")
names(colors) = c("Multiple of one type", "None", "Two of one type", "Two of two types")
necto_types_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(necto_types_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
```

@@

#####SIMMAP Habitat
```{r depth_zones, echo=F, message=F, warning=F}
#Simmap Vertical.Range
plotTree(ultratree, lwd = 4)
colors = c("black", "cyan", "blue", "green", "red")
names(colors) = c("Bathypelagic", "Surface", "Mesopelagic", "Epipelagic", "Deep benthic")
vrange_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(vrange_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
```

####Continuous traits: Brownian motion reconstructions

@@

#####Bathymetrics phylomorphospace
```{r depth_phylomorpho, echo=F, message=F, warning=F}
par(oma=rep(3,4))
phylomorphospace(depthmap$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=depthmap$cols, fsize=1, label = "horizontal")
```

#####Bathymetrics phylomorphospace with time from root as color
```{r depth_phylochronomorpho, echo=F, message=F, warning=F}
evotime = fastBM(dpruned_tree,nsim=2)
evotime_map <- contMap(dpruned_tree,evotime[,1], plot=F)
evotime_map$cols[]<-rainbow(1001,start=0.7,end=0)
phylomorphospace(evotime_map$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=evotime_map$cols, fsize=1, label = "horizontal")
#add.color.bar(2,evotime_map$cols,title="Time from the root", lims=c(0,2000),digits=5)
```

####Abundance of each taxon in the Monterey Bay sampling site
```{r counts_bars, echo=F, message=F, warning=F}
par(oma=rep(1,4))
log_abundance = log(abundance)
log_abundance["Physalia physalis"] = 0
plotBranchbyTrait(dpruned_tree, log_abundance, mode = "tips")
```

###Uncertainty inclusive traitgram for depth range
```{r uncertainty_depth, echo=F, message=F, warning=F}
contTrait <- -depth_range
A<-fastAnc(dpruned_tree,contTrait,CI=TRUE)
paintree<-paintSubTree(dpruned_tree,node=length(dpruned_tree$tip)+1,"1")
trans<-as.character(floor(0:50/2))
trans[as.numeric(trans)<10]<- paste("0", trans[as.numeric(trans)<10],sep="")
# plot
for(i in 0:50){
  p<-i/length(trans)
  phenogram(dpruned_tree,c(contTrait,(1-p)*A$CI95[,1]+p*A$ace), colors=setNames(paste("#0000ff",trans[i+1],sep=""),1), add=i>0)
  phenogram(dpruned_tree,c(contTrait,(1-p)*A$CI95[,2]+p*A$ace), colors=setNames(paste("#0000ff",trans[i+1],sep=""),1), add=TRUE)
}
phenogram(dpruned_tree,c(contTrait,A$ace),add=TRUE, colors=setNames("black",1))
```

####Bathymetrics scattergram array
```{r depth_scattergram, echo=F, message=F, warning=F}
fancyTree(dpruned_tree,type="scattergram",X=cbind(depth_max, depth_min, depth_range, depth_median),control=list(spin=FALSE))
```

####Distribution of sex across siphonophore taxa
```{r ggtree_sex, echo=F, message=F, warning=F, include=F}
exp_tree = ultratree
exp_tree$edge.length = 10*ultratree$edge.length
dd = data.frame(taxa  = tree$tip.label, trait = sex[match(tree$tip.label, names(sex))])
p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
p+theme(legend.position="right") + guides(color=guide_legend(title="Sex distribution", reverse=T)) 
```

####Tentilla presence across siphonophore taxa
```{r ggtree_tentillq, echo=F, message=F, warning=F, include=F}
exp_tree = ultratree
exp_tree$edge.length = 10*ultratree$edge.length
dd = data.frame(taxa  = tree$tip.label, trait = as.factor(tentilla[match(tree$tip.label, names(tentilla))]))
p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
p+theme(legend.position="right") + guides(color=guide_legend(title="Tentilla", reverse=T)) 
```

####Palpon presence across siphonophore taxa
```{r ggtree_palpon, echo=F, message=F, warning=F, include=F}
exp_tree = ultratree
exp_tree$edge.length = 10*ultratree$edge.length
dd = data.frame(taxa  = tree$tip.label, trait = as.factor(palpons[match(tree$tip.label, names(palpons))]))
p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
p+theme(legend.position="right") + guides(color=guide_legend(title="Palpons", reverse=T)) 
```

####Overview grid of main binary traits
```{r ggtree_bin_heatmap, echo=F, message=F, warning=F}
p <- ggtree(ultratree)
p <- p + geom_tiplab(size=2.2)
f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(1,2,5:8,12)]
rownames(f_cdata) = rownames(cdata)
f_cdata[match(tree$tip.label,rownames(f_cdata)),]
names(f_cdata)[5] = "Float"
gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 1)
```

####Phylogenetic signal in the character data

#####Agnostic branch length tree generation:
```{r flat_tree}
flat_tree = tree
flat_tree$edge.length = rep(mean(flat_tree$edge.length), length(flat_tree$edge.length))
```

```{r phylosig_binaries_declare, echo=F, message=F, warning=F, include=F}
binaries_phylosig = multiPhylosignal(as.matrix(cdata[,c(2,5:8)]), ultratree)
```
#####Phylogenetic signal in Binary Traits
```{r phylosig_binaries, echo=F, message=F, warning=F}
binaries_phylosig
```

```{r phylosig_depth_declare, message=F, warning=F, include=FALSE}
depth_physig = physignal(dpruned_data[,c(14,15,16,18,19)], dpruned_tree)
depth_multisig = multiPhylosignal(dpruned_data[,c(14,15,16,18,19)], dpruned_tree)
```
#####Phylogenetic signal in VARS bathymetrics
```{r phylosig_depth, echo=F, message=F, warning=F}
depth_physig
depth_multisig
```

#####Phylogenetic PCA for bathymetrics
```{r pPCA_depth, echo=F, message=F, warning=F}
phyl.pca(dpruned_tree, dpruned_data[,c(14,15,16,18,19)], mode="corr") %>% biplot

dpruned_data[,c(14,15,16,18,19)] %>% PCA(graph=F) -> Zpca
Zpca %>% fviz_screeplot(addlabels=T)
Zpca %>% fviz_contrib(choice="var", axes=1, sort.val="desc")
Zpca %>% fviz_pca_biplot( col.var="contrib", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), repel = TRUE)

phyl.pca(dpruned_tree, dpruned_data[,c(14,15,16,18,19)], mode="corr") -> ppcaZ
prcomp( dpruned_data[,c(14,15,16,18,19)]) -> rpcaZ
rpcaZ$rotation = ppcaZ$L
rpcaZ$x = ppcaZ$S
rpcaZ$scale = TRUE

rpcaZ %>% fviz_screeplot(addlabels=T)
rpcaZ %>% fviz_contrib(choice="var", axes=1, sort.val="desc")
rpcaZ %>% fviz_pca_biplot( col.var="contrib", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), repel = TRUE)

rpcaZ %>% fviz_pca_ind(repulse = T)
```

#####Phylogenetic ANOVA for depth
```{r phylANOVA_depth_declare, echo=F, message=F, warning=F, include=F}
z_nectopos = phylANOVA(dpruned_tree, dpruned_data$Nectosome.position, as.numeric(depth_median))
z_sex = phylANOVA(dpruned_tree, dpruned_data$Sex, as.numeric(depth_median))
z_necto = phylANOVA(dpruned_tree, dpruned_data$Nectosome, as.numeric(depth_median))
z_nectotype = phylANOVA(dpruned_tree, dpruned_data$Nectophores, as.numeric(depth_median))
z_palp = phylANOVA(dpruned_tree, dpruned_data$Palpons, as.numeric(depth_median))
z_tentilla = phylANOVA(dpruned_tree, dpruned_data$Tentilla, as.numeric(depth_median))
z_float = phylANOVA(dpruned_tree, dpruned_data$Pneumatophore, as.numeric(depth_median))
```
Each phylogenetic ANOVA compares variance of depth (continuous variable) within and between groups clustered by their categorical biological traits.

```{r phylANOVA_depth, echo=F, message=F, warning=F}
print("Depth vs Sex Distribution")
z_sex
print("Depth vs Nectosome")
z_necto
print("Depth vs Nectophore Types and Number")
z_nectotype
print("Depth vs Nectosome Position")
z_nectopos
print("Depth vs Palpons")
z_palp
print("Depth vs Tentilla")
z_tentilla
print("Depth vs Pneumatophore")
z_float
```

#####PGLS - Sex distribution
```{r pgls_depth_sex, echo=F, message=F, warning=F}
dp_sex <- dpruned_data$Sex
dp_sex = as.numeric(dp_sex)
names(dp_sex) = names(depth_median)
dp_sex[dp_sex==2] = 0        # 0 = Monoecious, 1 = Dioecious
#Vx <- Vy <- Cxy <- rep(0,length(dpruned_tree$tip.label))
#pgls.Ives(dpruned_tree, dp_sex, depth_median, Vx, Vy, Cxy)
pGLS = pgls.Ives(dpruned_tree, rep(dp_sex,3), rep(depth_median,3))
phylolm(depth_median~dp_sex,data=as.data.frame(cbind(depth_median,dp_sex)),phy=dpruned_tree,model="BM",measurement_error=TRUE,boot=100)
gls(dp_sex ~ depth_median)
```

@@

#Clean Raw Data
names(raw_data)[1] = "Species"
raw_data$Species[which(raw_data$Species == "Kephyes")] = "Kephyes ovata"
raw_data$Species[which(raw_data$Species == "Lilyopsis")] = "Lilyopsis fluoracantha"
raw_data$Species[which(raw_data$Species == "Agalma")] = "Agalma elegans"
raw_data$Species[which(raw_data$Species == "Frillagalma")] = "Frillagalma vityazi"
raw_data$Species[which(raw_data$Species == "Stephalia")] = "Stephalia dilata"
raw_data$Species[which(raw_data$Species == "Lychnagalma")] = "Lychnagalma utricularia"
raw_data$Species[which(raw_data$Species == "Craseoa")] = "Craseoa lathetica"
raw_data$Species[which(raw_data$Species == "Cordagalma")] = "Cordagalma ordinatum"
raw_data$Species[which(raw_data$Species == "Desmophyes haematogaster")] = "Prayidae D27SS7"
raw_data$Species[which(raw_data$Species == "Craseoa lathetica")] = "Prayidae D27D2"

@@

>>>>

```{r char_evo_setup, include=F, echo=F}
#Load data
read.csv('character_coding/main_characters.csv', header = T, sep = ',') -> cdata
cdata$Species = str_replace_all(cdata$Species, "sp ", "sp")
rownames(cdata) = cdata$Species

```{r simmap_making, echo=F, message=F, warning=F}
#Stochastic trait mapping:
#Make simmap objects
#print("SIMMAP Sex Distribution")
make.simmap(ultratree, sex, nsim = 100, message=FALSE) -> sex_sim
#print("SIMMAP Palpons")
make.simmap(ultratree, palpons, nsim = 100, message=FALSE) -> palpons_sim
#print("SIMMAP Tentilla")
make.simmap(ultratree, tentilla, nsim = 100, message=FALSE) -> tentilla_sim
#print("SIMMAP Tentilla2")
make.simmap(ultratree, tentilla2, nsim = 100, message=FALSE) -> tentilla_sim2
#print("SIMMAP Nectosome")
make.simmap(ultratree, nectosome, nsim = 100, message=FALSE) -> nectosome_sim
#print("SIMMAP Nectosome Position")
make.simmap(ultratree, necto_pos, nsim = 100, message=FALSE) -> necto_pos_sim
#print("SIMMAP Nectophore Types")
make.simmap(ultratree, necto_types, nsim = 100, message=FALSE) -> necto_types_sim
#print("SIMMAP Bracts")
make.simmap(ultratree, bracts, nsim = 100) -> bracts_sim
#print("SIMMAP Habitat")
make.simmap(ultratree, vrange, nsim = 100, message=FALSE) -> vrange_sim

@@

@@

@@

@@

@@

####Overview grid of main binary traits
```{r ggtree_bin_heatmap, echo=F, message=F, warning=F}
p <- ggtree(ultratree)
p <- p + geom_tiplab(size=2.2)
f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(1,2,5:8,12)]
rownames(f_cdata) = rownames(cdata)
f_cdata = f_cdata[,c(4,7,1,3,6,2,5)]
f_cdata[match(tree$tip.label,rownames(f_cdata)),]
gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 1)
```

####Phylogenetic signal in the character data
```{r phylosig_binaries_declare, echo=F, message=F, warning=F, include=F}
binaries_phylosig = multiPhylosignal(as.matrix(cdata[,c(2,5:8)]), ultratree)
```

#####Phylogenetic signal in Binary Traits
```{r phylosig_binaries, echo=F, message=F, warning=F}
binaries_phylosig # make it xtable
```

#####Phylogenetic signal in VARS bathymetrics
```{r phylosig_depth, echo=F, message=F, warning=F}
depth_physig
depth_multisig
```

@@

#Clean Raw Data
names(raw_data)[1] = "Species"
raw_data$Species = as.character(raw_data$Species)
raw_data$Species[which(raw_data$Species == "Kephyes")] = "Kephyes ovata"
raw_data$Species[which(raw_data$Species == "Lilyopsis")] = "Lilyopsis fluoracantha"
raw_data$Species[which(raw_data$Species == "Agalma")] = "Agalma elegans"
raw_data$Species[which(raw_data$Species == "Frillagalma")] = "Frillagalma vityazi"
raw_data$Species[which(raw_data$Species == "Stephalia")] = "Stephalia dilata"
raw_data$Species[which(raw_data$Species == "Lychnagalma")] = "Lychnagalma utricularia"
raw_data$Species[which(raw_data$Species == "Craseoa")] = "Craseoa lathetica"
raw_data$Species[which(raw_data$Species == "Cordagalma")] = "Cordagalma sp"
raw_data$Species[which(raw_data$Species == "Desmophyes haematogaster")] = "Prayidae D27SS7"
raw_data$Species[which(raw_data$Species == "Craseoa lathetica")] = "Prayidae D27D2"
raw_data$Species[which(raw_data$Species == "Rudjakovia")] = "Rudjakovia sp"
raw_data$Species[which(raw_data$Species == "Thermopalia")] = "Thermopalia sp"
raw_data$Species[which(raw_data$Species == "Physonectae sp. 1")] = "Physonect sp"

@@

#90% Cut-off of extrame values
SPDepths = parched_data[,1:2]
byspp_SPD = split(SPDepths, SPDepths$Species)
for(i in 1:length(byspp_SPD)){
  if(nrow(byspp_SPD[[i]])>200){
#   byspp_SPD[[i]] = sample_n(byspp_SPD[[i]], 0.02*nrow(byspp_SPD[[i]]))
    byspp_SPD[[i]] = sample_n(byspp_SPD[[i]], 200)
  }
}
trimmed_SDdata = bind_rows(byspp_SPD)

sp_order = c("Agalma elegans", "Athorybia rosacea", "Halistemma rubrum", "Nanomia bijuga", "Lychnagalma utricularia", "Physophora gilmeri", "Forskalia asymmetrica", "Cordagalma sp", "Resomia ornicephala", "Frillagalma vityazi", "Physonect sp", "Marrus claudanielis", "Stephalia dilata", "Thermopalia sp", "Erenna richardi", "Abylopsis tetragona", "Diphyes dispar", "Chelophyes appendiculata", "Kephyes ovata", "Chuniphyes multidentata", "Lilyopsis fluoracantha", "Prayidae D27D2", "Prayidae D27SS7", "Hippopodius hippopus", "Bargmannia amoena", "Bargmannia elongata", "Bargmannia lata", "Rudjakovia sp", "Apolemia rubriversa", "Apolemia lanosa", "Apolemia sp", "Rhizophysa filiformis", "Physalia physalis")

trimmed_SDdata$Species = as.character(trimmed_SDdata$Species)
trimmed_SDdata = trimmed_SDdata[which(!is.na(trimmed_SDdata$Species)),]
trimmed_SDdata$Species <- factor(trimmed_SDdata$Species, levels = sp_order)

ggplot(trimmed_SDdata, aes(x = Species, y=-Depth)) + geom_boxplot() + theme_classic()
ggplot(trimmed_SDdata, aes(x = Species, y=-Depth)) + geom_violin() + theme_classic()
